---
title: "HM2"
author: "by Tamires Amorim"
date: "3/2/2021"
output: html_document
---

Loading packages:

```{r library, echo=FALSE}
library(tidyverse)
library(data.table)
library(DT)
library(dplyr)
```

Understanding what is the information contained in the tables, before starting the analysis: 
```{r}
glimpse(table1) #identification of characters, integers and so on.

# on the console write: ?table1 , there I found the description of the data in all the tables

#all display the number of TB cases documented by the World Health Organization in Afghanistan, Brazil, and China between 1999 and 2000. The data contains values associated with four variables (country, year, cases, and population), but each table organizes the values in a different layout.

```

### Homework 2:

#### 12.2.1 Exercises 

##### Using prose, describe how the variables and observations are organised in each of the sample tables
Sample table 1: The rows have country, and year. The columns cases and populations related to the year and country.
Sample table 2: Each row describes a country and year combined. The column "count" includes cases and population in separated rows.
Sample table 3: the rows represent country and year, the columns shows the rate (number of cases divided by population).
Sample table4a: shows values of cases, where the row is the country and the column is the year (1999,2000).
sample table4b: represents values of population, where the row is the country and the column is the year (1999,2000).
Sample table 5: works similarly with table 3, the difference is that there is an addition of the column century.

```{r}
table1
table2
table3
table4a
table4b
table5
```
##### Compute the rate for table2, and table4a + table4b. You will need to perform four operations:
From my perspective, the table 2 is harder because it has separate rows for cases and population, which lead me to create a new table with the columns for calculating cases per capita. For table4a and 4b since it is already split, meaning one for cases and the other for population, I just had to divide cases by population. 

Table 2:
```{r}
tb2_cases <- filter(table2, type == "cases") %>%
  rename(cases = count) %>%
  arrange(country, year)
tb2_population <- filter(table2, type == "population") %>%
  rename(population = count) %>%
  arrange(country, year)

###checking if it worked:

tb2_cases
tb2_population
```

```{r}
## data frame with population and cases columns 

tb2_cases_per_cap <- tibble(
  year = tb2_cases$year,
  country = tb2_cases$country,
  cases = tb2_cases$cases,
  population = tb2_population$population
) %>%
  mutate(cases_per_cap = (cases / population) * 10000) %>%
  select(country, year, cases_per_cap)

tb2_cases_per_cap
```

```{r}
## Storing the variable as a new row in the table two

tb2_cases_per_cap <- tb2_cases_per_cap %>%
  mutate(type = "cases_per_cap") %>%
  rename(count = cases_per_cap)

bind_rows(table2, tb2_cases_per_cap) %>%
  arrange(country, year, type, count)
```


Table 4a and 4b:
```{r}
table4_cases_per_capita <- tibble(
    country = table4a$country,
    `1999` = table4a[["1999"]] / table4b[["1999"]] * 10000,
    `2000` = table4a[["2000"]] / table4b[["2000"]] * 10000
  )
table4_cases_per_capita
```

##### Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?

First we need to filter the table2 with rows that represent cases of Tuberculosis.Then plot the relationship between cases by year and country. 
```{r}
table2 %>%
  filter(type == "cases") %>%
  ggplot(aes(year, count)) +
  geom_line(aes(group = country), colour = "grey50") +
  geom_point(aes(colour = country)) +
  scale_x_continuous(breaks = unique(table2$year)) +
  ylab("cases")
```

#### b) 12.3.3 Exercises

##### Why are pivot_longer() and pivot_wider() not perfectly symmetrical? Carefully consider the following example:
They are not symmetrical because pivot_longer() makes wide tables narrower and longer; pivot_wider() makes long tables shorter and wider. When converting from wide to long you can lost the information from column type. 

The names_to="", in the pivot_longer() function, makes the column to be a character (name).

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
    pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return")
```
#### c) 12.5.1 Exercises

##### Compare and contrast the fill arguments to pivot_wider() and complete().
To replace NAs in the data set you can add fill or complete to the pivot_wider() function. The values_fill argument in pivot_wider() "takes the set of columns where you want missing values to be replaced by the most recent non-missing value". the complete() "takes a set of columns, and finds all unique combinations. It then ensure the original dataset contains all those values, filling in explicit NAs where necessary". The difference is that values_fill accepts a single value, and complete() it can have sets of values to replace NAs.

For Values_fill = 0, we have:
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)
stocks %>% 
  pivot_wider(names_from = year, values_from = return,
              values_fill = 0)
```
For the Complete() argument:
```{r}
stocks %>% 
 complete(year, qtr, fill=list(return=0))
```

##### What does the direction argument to fill() do?

The direction will determine if NA values should be replaced by the previous non-missing value ("down") or the next non-missing value ("up").
```{r}
# No fill
stocks

# fill with direction = "down"
stocks %>% fill(return, .direction = "down")


# fill with direction = "up"
stocks %>% fill(return, .direction = "up")
```



